---
title: "320FinalProject"
authors: "Charles Bangert, Noah Tanenholtz, Ryan Dietz"
date: "5/9/2019"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Final Project for CMSC320: Introduction to Data Science

# Summary
In lieu of a final exam, CMSC320 students will turn in a tutorial that will walk users through the entire data science pipeline: data curation, parsing, and management; exploratory data analysis; hypothesis testing and machine learning to provide analysis; and then the curation of a message or messages covering insights learned during the tutorial. Students may choose an application area and dataset(s) that are of interest to them; please feel free to be creative about this!

Remember, the course resources page has a number of data repositories from where you can download data. You could also create your own dataset using scraping skills you have learned in class. We in fact, encourage this.

The tutorial should be self-contained as an Rmarkdown document, and delivered as a GitHub statically-hosted Page (described below). You can see examples here: https://www.kaggle.com/kernels?sortBy=votes&group=everyone&pageSize=20&language=R

In general, the tutorial should contain at least 1500 words of prose and 150 lines of (nonpadded, legitimate) R code, along with appropriate documentation, visualization, and links to any external information that might help the reader.

#Get data 
 
 May 9, 2018 - May 9, 2019
```{r}
library(tidyverse) ;
library(httr) ; 
```
```{r data}
library(tidyverse)
library(httr)  

df

csv <- read_csv("pergame_data.csv")  
pergame <- csv %>% data.frame()
pergame

csv1 <- read_csv("totals_data.csv")
totals <- csv1 %>% data.frame()
totals

csv2 <- read_csv("per36min.csv")
per36min <- csv1 %>% data.frame()
per36min

csv3 <- read_csv("per100poss.csv")
per100poss <- csv3 %>% data.frame()
per100poss

csv4 <- read_csv("advanced.csv")
advanced <- csv4 %>% data.frame()
advanced


```
- *****


#Basic Operations

- To make the data column names more recognizable of what they mean, we are going to use some basic R operations to change the column names and even delete columns as a whole.

```{r basics}

names(pergame) <- c("rank", "player", "year_from", "year_to", "games_played", "games_stared", "min_played", "fg_made", "fg_attempt", "fg_percent", "3p_made", "3p_attempt", "3p_percent", "2p_made", "2p_attempt", "2p_percent", "eff_fg_percent", "ft_att", "ft_made", "ft_percent", "off_rebounds", "def_rebounds", "total_rebounds", "assists", "steals", "blocks", "turnovers", "fouls", "points")

#This way of changing column names is very easy and useful if you are going to change all the column names. You start with the first column and go all the to the last column
pergame

colnames(totals)[colnames(totals) == "Rk"] <- "rank"
colnames(totals)[colnames(totals) == "Player"] <- "player"
colnames(totals)[colnames(totals) == "From"] <- "year_from"
colnames(totals)[colnames(totals) == "To"] <- "year_to"
colnames(totals)[colnames(totals) == "G"] <- "games_played"
colnames(totals)[colnames(totals) == "GS"] <- "games_started"
colnames(totals)[colnames(totals) == "MP"] <- "min_played"
colnames(totals)[colnames(totals) == "FG"] <- "fg_made"
colnames(totals)[colnames(totals) == "FGA"] <- "fg_attempt"
colnames(totals)[colnames(totals) == "FG."] <- "fg_percent"
colnames(totals)[colnames(totals) == "X3P"] <- "3p_made"
colnames(totals)[colnames(totals) == "X3PA"] <- "3p_attempt"
colnames(totals)[colnames(totals) == "X3P."] <- "3p_percent"
colnames(totals)[colnames(totals) == "X2P"] <- "2p_made"
colnames(totals)[colnames(totals) == "X2PA"] <- "2p_attempt"
colnames(totals)[colnames(totals) == "X2P."] <- "2p_percent"
colnames(totals)[colnames(totals) == "eFG."] <- "eff_fg_percent"
colnames(totals)[colnames(totals) == "FT"] <- "ft_made"
colnames(totals)[colnames(totals) == "FTA"] <- "ft_attempt"
colnames(totals)[colnames(totals) == "FT."] <- "ft_percent"
colnames(totals)[colnames(totals) == "ORB"] <- "off_rebounds"
colnames(totals)[colnames(totals) == "DRB"] <- "def_rebounds"
colnames(totals)[colnames(totals) == "TRB"] <- "total_rebounds"
colnames(totals)[colnames(totals) == "AST"] <- "assists"
colnames(totals)[colnames(totals) == "STL"] <- "steals"
colnames(totals)[colnames(totals) == "BLK"] <- "blocks"
colnames(totals)[colnames(totals) == "TOV"] <- "turnovers"
colnames(totals)[colnames(totals) == "PF"] <- "fouls"
colnames(totals)[colnames(totals) == "POints"] <- "points"

#This way of changing  a column name is very easy and useful if you are going to change only a few of the column names. You can change the names in any order. It is very repetive and unefficient way to change column names if you are changing all the names as we did.
totals

names(per36min) <- c("rank", "player", "year_from", "year_to", "games_played", "games_stared", "min_played", "fg_made", "fg_attempt", "fg_percent", "3p_made", "3p_attempt", "3p_percent", "2p_made", "2p_attempt", "2p_percent", "eff_fg_percent", "ft_att", "ft_made", "ft_percent", "off_rebounds", "def_rebounds", "total_rebounds", "assists", "steals", "blocks", "turnovers", "fouls", "points")
per36min

#If you looked over all the data that was read in, in the previous step, then you should of noticed a blank column in the dataset of per100poss and advanced. We are going to show the steps needed to delete the blank columns and then change the column names for the two datasets

#We see that the error column in per100poss is named X29, you are going to need to know the column name to delete it. To delete the column, we set the error column to NULL

per100poss$X29 <- NULL

names(per100poss) <- c("rank", "player", "year_from", "year_to", "games_played", "games_stared", "min_played", "fg_made", "fg_attempt", "fg_percent", "3p_made", "3p_attempt", "3p_percent", "2p_made", "2p_attempt", "2p_percent", "ft_att", "ft_made", "ft_percent", "off_rebounds", "def_rebounds", "total_rebounds", "assists", "steals", "blocks", "turnovers", "fouls", "points", "off_rating", "def_rating")
per100poss

#We see that the error column in advanced is named X19 and X24

advanced$X19 <- NULL
advanced$X24 <- NULL
 
names(advanced) <- c("rank", "player", "year_from", "year_to", "games_played", "min_played", "eff_rating", "true_shooting_percent", "3p_attempt_rate", "ft_attempt_rate", "off_rebound_percent", "def_rebound_percent", "total_rebound_percent", "assist_percent", "steal_percent", "block_percent", "turnover_percent", "usage_percent", "off_win_share", "def_win_share", "win_share", "ws_per48min", "off_plus_minus", "def_plus_minus", "plus_minus", "value_over_replace")
advanced

```



